Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-20 paper.qmd
Add sentence to mention about video links
Updated Figure 10 title

The paper is organized as follows. @sec-background provides a summary of the literature on NLDR, and high-dimensional data visualization methods. @sec-method contains the details of the new methodology, including simulated data examples. In @sec-bestfit, we describe how to assess the best fit and identify the most accurate \gD{} layout based on the proposed model diagnostics. Curiosities and unexpected patterns discovered in NLDR results by examining the model in the data space are discussed in @sec-curiosities. Two applications illustrating the use of the new methodology for bioinformatics and image classification are in @sec-applications. Limitations and future directions are provided in @sec-discussion.

@@

```{r}
#| label: fig-one-dens_clust-error
#| fig-width: 15
#| fig-height: 10
#| fig-pos: H
#| out-width: 100%
#| fig-cap: "Exploring the effect of density on the NLDR layout using a C-shaped structure with different density in each end. The sparse end of the structure shows a high model error, while the dense end shows a low model error. The top part of the triangular shaped with linked brushing. The tSNE layout, with dense points (a1-a3) are colored according to their \\fD{} model error. Darker colors represent a high error, while lighter colors indicate a low error. a2, a3 and b2, b3 present the results from linked plots that brushed the high \\fD{} error points in \\gD{} and \\fD{}. It helps in identifying the high \\fD{} model errors that occur due to the sparse end."

>>>>

The paper is organized as follows. @sec-background provides a summary of the literature on NLDR, and high-dimensional data visualization methods. @sec-method contains the details of the new methodology, including simulated data examples. In @sec-bestfit, we describe how to assess the best fit and identify the most accurate \gD{} layout based on the proposed model diagnostics. Curiosities and unexpected patterns discovered in NLDR results by examining the model in the data space are discussed in @sec-curiosities. Two applications illustrating the use of the new methodology for bioinformatics and image classification are in @sec-applications. Limitations and future directions are provided in @sec-discussion. Links to the langevitour animation videos showing the \gD{} projections are provided in @tbl-links.

@@

```{r}
#| label: fig-one-dens_clust-error
#| fig-width: 15
#| fig-height: 10
#| fig-pos: H
#| out-width: 100%
#| fig-cap: "Exploring the effect of density on the NLDR layout using a C-shaped structure with different density at each end. The top part of the triangular shape, shown with linked brushing, identify high model error corresponding to the sparse end of the structure (a1-a3). On the other hand, one of the dense corners of the structure shows low model error and is highlighted in the right part of the triangular shape (b1-b3). The tSNE layouts (a1, b1) are colored according to selected points. a2, a3 and b2, b3 present the results from linked plots that brushed the \\fD{} error points in \\gD{} and \\fD{}. It helps in identifying the high \\fD{} model errors that occur due to the sparse end."
